---
title: "lab-07-simpsons.Rmd"
author: "yazi mohammad "
date: "19 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
view(Whickham)
```
    
Your answer:
The date is observational as the description states that is based on age, smoking , and mortality , which are all observable events and not produced via experiments.

2.  
```{r observations}
nrow(Whickham)
```

Your answer;
There are 1.314 observations. As we know every row is an observation.

3.  

```{r number_variables}
 names(Whickham)
```

Your answer:  
There are 3 variables, "outcome" , "smoker" , and "age"


```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)
```

Your answer: 
Using the `unique()` function on the 3 variables we could see that "outcome" only takes Alive or Dead value, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age numberical continous date.

One of the best ways to visualise categorical date is through the use of bar charts.

```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +
  geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +
  geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +
  geom_boxplot()
```


4. 
I expect the health will be worser and may the person will be died after while , if he keeping smoke

```{r plot_smokeVSincome_1}
ggplot(Whickham, aes(x=smoker, y=outcome , color=outcome)) +
  geom_bar(stat = "identity")
```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
```

6.

```{r age_cat}

```

7. 

```{r plot_smokeVSincome_3}

```

Knit, commit, and push to github.